### 2021-03-13 chapters/sections/1-context.tex
Anfang und ende fertig

Since invention of the World Wide Web, the number of its users increased rapidly. Today, more than 90 percent of the german population uses the internet \cite{Onlinestudie}. Enterprises and organizations recognised this potential to provide their services to a large number of people as Service Providers (SP). A frequently used method of SPs is "online self-service" (OSS). Specialized software tools enable customers to use services through the internet without direct human interaction.

Many tools require identification of a user in case interactions or initiated business processes have to be associated with him. A system for management of identities of users therefore is part of numerous system architectures. Service Providers usually enable users to manage their partial identities using online self-service through user profiles. Today, customers of SPs are associated with multiple partial identities and user profiles - one for each SP.

The result is a variety of problems in management of digital identities \cite{IdentityCrisis}. Here are some examples:
\begin{itemize}
    \item Multiple SPs collect, use and share parts of identities which leads to distributed and fragmented identities.
    \item As a result of the distributed identities, risk of data breaches increases.
    \item Managing multiple user profiles is very inconvenient for users.
    \item It is often unclear who stores which data and for what purpose.
    \item SPs have an incentive to collect and share more user data than they need.
\end{itemize}

An improved model for identity management is necessary. In recent years, multiple new models have been invented. One model called Social Login became popular. It enables customers to create User Profiles for SPs through existing User Profiles of Social Networks like Facebook. This improves identity management by enabling SPs to rely on the external profile for unique identification, authorisation and personal attributes. Users can therefore manage identity related information through their Social Profile which is accessed by the SP.

This approach, however, still leaves the requirement of User Profiles for each SP because, depending on the provided service, additional attributes which are not part of the Social Profile, have to be manageable by the user. SPs also might require additional online self-service tools the social network does not provide.

The solution to this problem is an Identity Management Provider (IMP) which manages the whole online identity of a person and can replace the individual partial identities and user profiles. The system itself is provided and hosted by a SP for consumption by other SPs. To replace existing User Profiles, the IMP can not only rely on systems for registration, authentication, authorisation and provisioning. Additional requirements of SPs like communication, data wallet and management of additional attributes have to be satisfied.

In order for an SP to use identities and services of an IMP, an integration into the system architecture of the SP has to take place. Due to the complex nature of system architectures and identity management, this is a difficult task.

A currently relevant example for usage of online self-service with particular interesting requirements for identities is the German "Online Access Law" (Online Zugangs Gesetz: OZG). It requires all administrative services of the german federal republic, each member state and commune to be digitally available through interoperable user profiles. The current plan is to make the profiles only available for usage in context of the OZG. From a user perspective, this would be yet another partial identity and user profile to manage.

>>>>

Many services, are accessible through the internet. Especially during the COVID-19 pandemic, stores had to switch to online trade due to lockdowns. Online Service Providers (SP) often require customers to create user profiles in order to access their services. Management of personal information as well as interaction and communication between Service Provider and customer are often only possible through a web page or e-mail. It is time consuming to maintain even a small amount of user profiles: To change personal information, like an e-mail address, the web page of each Service Provider has to be accessed the correct credentials have to be submitted and the information has to be manually filled in each time.

In some cases, user profiles are required by Service Providers even if their service could be provided without them. Especially in case of small stores, running their individual online shop, user profiles are not necessary for customers to place an order. Personal information can be shared temporarily until processing of the order is completed and the package delivered. Some online shops allow users to place orders by temporarily sharing personal information, which however, limits interaction and communication possibilities to insecure e-mail exchange.

A method for identity management called Identity Management Provisioning (IMP) is introduced. IMP provides identity management as a service to users and Service Providers. It enables users to create and manage a single digital identity and use it to share personal information and interact with multiple Service Providers through "IMP relationships" and "IMP messages". Through IMP relationships, users and Service Providers have a tool to securely share personal information which, depending on the use case can be used in different ways. It can be used to create and manage a persistent user profile or to temporarily share personal information for a business process. IMP messages enable user and Service Provider to securely exchange structured data and depending on the client application which receives and displays the data, design rich user interactions.

### 2021-03-13 chapters/sections/10-imp_solution_integration.tex
Anfang und ende fertig

The IMP solution presented in this section describes the utilization of IMP relationships with the purpose of creating and maintaining user profiles through the IMP client. Message types for attribute synchronization, authentication and communication are defined. To not disrupt the operation of the existing system architecture, the IMP solution is designed to not replace user profiles but provide an additional way of accessing them trough IMP. To later reduce technological integration efforts, the IMP solution is designed to only require interaction with systems of the portal domain.

@@

\begin{itemize}
    \item Title: Create OZG User Profile
    \item Attributes: Name, Surname, ... 
    \begin{itemize}
        \item Name: Max
        \item Surname: Mustermann
    \end{itemize}
    \item Shared Attributes: Administration Portal of the member state ...
    \item Reason: Create an OZG user profile and connect your IMP identity.
    \item Metadata: <<data for technological integration>>
\end{itemize}

Users visiting the account creation web page of the administration portal can use the IMP client to scan the QR code. The IMP client presents the corresponding relationship template to the user and automatically fills in attributes which the administration portal requests from the user as part of the relationship. It is assumed that the IMP system has systems in place for identifying users and validating personal information of corresponding IMP identities. As a result of that, service providers could rely on the validity of shared attributes.

@@

\subsection{Message Utilization}
As part of an established relationship, service provider and user can securely exchange structured data in form of asynchronous messages. Security in this case means that messages can not be accessed or modified by a third party and are authorized by the corresponding user. Based on this secure data exchange capabilities, service providers can define various types of messages, each with different purposes and effects on the IMP client. This section contains various message types useful for managing the user profile created as a result of an established relationship.

\paragraph{Login to User Profile}
During the user profile creation process using IMP, the user specified a username but not a password. The IMP client is used for authentication of login requests to the administration portal.

@@

The IMP client is used for management of personal information of the user profile. Through the IMP client, the user can send requests for adding, updating or deleting shared attributes. Administration portal and IMP client exchange attribute change request and reply messages. The requests can either be accepted or rejected by the service provider. Accepting the request will result in according modification of the user profile connected to the relationship.

For simplification purposes, the web page for management of personal information on the administration portal is expected to be locked. However, it would also be possible for the portal to send attribute change requests to the IMP client based on the modifications the user performed on the web page.

@@

User profiles are known to the existing system architecture and are an often used identity management solution. Many service providers, especially in the private sector will want to keep using a profile based identity management approach. The IMP solution presents an opportunity of leveraging existing user profile systems while simplifying identity management through IMP. In addition to that, the IMP solution is designed, so that it can almost effortlessly be turned on and off, giving service providers the freedom of testing the solution. Service providers can initially add the IMP solution as an alternative to the conventional user profile identity management solution and after some period of time switch to only using the IMP solution.

Mapping user profiles to IMP relationships has many of the usability advantages described in section 3.2. The onboarding process is simplified. Users don't have to manually fill in a form for the creation of a user profile but are supported by the IMP client. As part of established relationships, service provider and users are able to directly exchange messages with structured content. In addition to interaction capabilities of the IMP client, this enables a variety of interaction scenarios not possible through for example common e-mail clients. In this case, scenarios for attribute changes, user profile authentication and communication are made possible.

This approach also increases data protection by providing the user with information on exactly which personal information is shared and how it is being processed as part of the profile creation process. This helps users in making an informed decision on whether to create the user profile. At any time after the onboarding process, users can view the relationship in the IMP client and check which service providers process which personal information and how. This enables users to keep an overview of the distribution of personal information.

Security is also increased as a result of this approach. Through established relationships, service provider and users have the ability to not only directly exchange messages but also securely. As a result of that, authentication through the IMP client can replace password based authentication methods. This eliminates the risk of weak or stolen passwords.

@@

After an effortless onboarding process on multiple administration portals, users are able to manage each profile through the IMP client. This enables users break free access to all user profiles giving the illusion of an interoperable user profile. Logging in to a user profile using the IMP client eliminates the need of remembering passwords, furthering the illusion of interoperable user profiles. Finally, if for each user profile, the same profile name is selected, the user wont be able to distinguish user profiles most of the time.

@@

The IMP solution has also several problems regarding usability and data protection.

Most importantly, as a result of integrating the IMP system through user profiles, each administration portal is still required to access personal information.

In addition to that, application and management of administrative services remains accessible only through administration portal and form server which presents a discontinuity in accessing OZG services.

>>>>

The IMP solution presented in this section describes the utilization of IMP relationships with the purpose of creating and maintaining user profiles through the IMP client. Message types for attribute synchronization, authentication and communication are defined. To not disrupt the operation of the existing system architecture, the IMP solution is designed to not replace user profiles but to provide an additional way of accessing them through IMP. To later reduce technological integration efforts, the IMP solution is designed to only require interaction with systems of the portal domain.

@@

\begin{figure}[H]
\begin{itemize}
    \item Title: Create OZG User Profile
    \item Attributes: Name, Surname, ... 
    \item Shared Attributes: Administration Portal of the member state ...
    \item Reason: Create an OZG user profile and connect the IMP identity.
    \item Metadata: <<data for technological integration>>
\end{itemize}
\end{figure}

Users visiting the account creation web page of the administration portal can use the IMP client to scan the QR code. The IMP client presents the corresponding relationship template to the user and automatically fills in attributes that the administration portal requests from the user as part of the relationship. It is assumed that the IMP system has systems in place for identifying users and validating personal information of corresponding IMP identities. As a result of that, Service Providers can rely on the validity of shared attributes.

@@

\subsection{Message Utilization}
As part of an established relationship, Service Providers and users can securely exchange structured data in form of asynchronous messages. Security in this case is defined as that a messages can not be accessed or modified by a third party and are authorized by the corresponding user. Based on this secure data exchange capabilities, Service Providers can define various types of messages, each with different purposes and effects on the IMP client. This section contains various message types useful for managing the user profile created as a result of an established relationship.

\paragraph{Login to User Profile}
During the user profile creation process using IMP, the user specifies a username but not a password. The IMP client is used for authentication of login requests to the administration portal.

@@

The IMP client is used for management of personal information of the user profile. Through the IMP client, the user can send requests for adding, updating or deleting shared attributes. Administration portal and IMP client exchange attribute change request and reply messages. The requests can either be accepted or rejected by the Service Provider. Accepting the request will result in according modification of the user profile connected to the relationship.

For simplification purposes, the web page for the management of personal information on the administration portal is expected to be read-only. However, it would also be possible for the portal to send attribute change requests to the IMP client based on the modifications the user performed on the web page.

@@

User profiles are known to the existing system architecture and are an often used identity management solution. Many Service Providers, especially in the private sector will want to keep using a profile based identity management approach. The IMP solution presents an opportunity of leveraging existing user profile systems while simplifying identity management through IMP. In addition to that, the IMP solution is designed, so that it can almost effortlessly be turned on and off, giving Service Providers the freedom of testing the solution. Service Providers can initially add the IMP solution as an alternative to the conventional user profile identity management solution and after a certain period of time switch to only using the IMP solution.

Mapping user profiles to IMP relationships has many of the usability advantages described in section 3.2. The onboarding process is simplified. Users don't have to manually fill in a form for the creation of a user profile but are supported by the IMP client. As part of established relationships, Service Provider and users are able to directly exchange messages with structured content. In addition to interaction capabilities of the IMP client, this enables a variety of interaction scenarios not possible through for example common e-mail clients. In this case, scenarios for attribute changes, user profile authentication and communication are made possible.

This approach also increases data protection by providing the user with information on exactly which personal information is shared and how it is being processed as part of the profile creation process. This helps users in making an informed decision on whether to create the user profile. At any time after the onboarding process, users can view the relationship in the IMP client and check which Service Providers process which personal information, why and how. This enables users to keep an overview of the distribution of their personal information.

Security is also increased as a result of this approach. Through established relationships, Service Provider and users have the ability to not only directly exchange messages but also to do this securely. As a result of that, authentication through the IMP client can replace password based authentication methods. This eliminates the risk of weak or stolen passwords.

@@

After an effortless onboarding process on multiple administration portals, users are able to manage each profile through the IMP client. This enables users to break free access from all user profiles giving the illusion of an interoperable user profile. Logging in to a user profile using the IMP client eliminates the need of remembering passwords, furthering the illusion of interoperable user profiles. Finally, if for each user profile, the same profile name is selected, the user wont be able to distinguish user profiles most of the time.

@@

The IMP solution has also several problems regarding usability and data protection. Most importantly, as a result of integrating the IMP system through user profiles, each administration portal is still required to access personal information. In addition to that, application and management of administrative services remains accessible only through the administration portal and form server which presents a discontinuity in accessing OZG services.
